Predicting Tags for StackOverflow Posts

### 2013-02-26 stackExchange/writeups/iccm-2013/writeup.tex
foo
Former-commit-id: 116c33b684092a62fa118d2803c083a0d275c64f [formerly cfc0a3e7d56f54495f8848cddb814753432602d9]
Former-commit-id: 2685ecfe70f9c2f40ce11e521a6c5fe4777d2c23
foo
Former-commit-id: 59b7539dd1c5d85f229c5c2e4f9d2c4f4ac15d25 [formerly 23eabfa0fd685faaca2f94633c9e4e0fbdf185a9]
Former-commit-id: bb1bd5d62f30f31e88df90e00e119f531810ce1c

\usepackage{cogsci}
\usepackage{pslatex}
\usepackage{apacite}
\usepackage[none]{hyphenat}

@@

Human-based tagging of online content has increased in recent years.
A few examples of human-based tagging of information are hashtags for Twitter tweets and Facebook posts, as well as keywords for research papers.
A large portion of the growing number of social media sites support some sort of human-directed tagging of posts.
Additionally, an increasing number of individuals are using tags as a form of query-based search to find information
(e.g., \citeNP{Diakopoulos2010}).

There is a growing body of research that aims to characterize the lifetime and growth characteristics of hashtags for real-time social networking sites such as Twitter
(e.g., \citeNP{Tsur2012, Chang2010}).
However, less is known about the underlying process when choosing particular tags during the creation of the tweet or post.
Modeling tag creation accurately (as opposed to modeling tag growth and behavior after creation) could allow a system to predict the hashtag that a user should use, given the contextual clues available.
There are numerous benefits to this type of tag prediction.
For example, users could be introduced to tags (and potentially communities surrounding those tags) that are of interest to them.
Also tag cleanup systems could be implemented where mistags are identified and corrected.

@@

@@

\subsubsection{Distribution of Tag Frequency}

\emph{C\#} is the most often used tag, where about 11\% of posts were tagged with \emph{C\#}.
Tag frequency then drops off sharply.
The top 1\% of tags based on frequency of use account for approximately 60\% of total tag instances.
This is why accounting for tag frequency (by including ACT-R's base-level activation component) is so important for the StackOverflow dataset.

@@

During initial model testing, the attentional weighting for all words was equal, and set to $W/n$ for $n$ title words (and set similarly for body words).
The problem with this approach is that familiar stop words like ``the'' and ``?'', that are often observed across all tags have the same attentional weighting as highly predictive words like ``PHP''.
Granted the $S_{ji}$'s across all tags for context words like ``the'' are much lower (and more evenly distributed) than the few $S_{ji}$'s that spike for ``PHP'' (e.g., the tag \emph{PHP}).
However, this still seemed problematic for two reasons:
First, because stop words like ``the'' are observed with all tags and not highly predictive of any tag, one could argue that their presence is simply adding error in the model's tag activation calculation.
Second, because $W$ is held constant across posts, any attentional resources that are allocated to stop words like ``the'' and ``?'' are taken away from being used for highly predictive words like ``PHP''.

@@

@@

Model tag predictions are based on the amount of word co-occurrences between words in the title and body of the post and tags used for that post.
To build the co-occurrence matrix between post words and tags, all title and body text and associated tags were extracted from the relational database.
The Python Natural Language Processing Toolkit \cite{Bird2009} was used to chunk (i.e., tokenize and lemmatize) the text in the post.
The tags were already chunked, but tag synonyms were converted to their root tag using the community-maintained StackOverflow tag synonym database.
Afterwords the processed word chunks and root tags were imported into a PostgreSQL database table.

@@

@@

>>>>

\usepackage{cogsci}
\usepackage{pslatex}
\usepackage{apacite}
\usepackage[none]{hyphenat}
\usepackage[compact]{titlesec}

@@

Human-based tagging of online content has increased in recent years.
A few examples are authors creating hashtags for Twitter tweets and Facebook posts, as well as keywords for research papers.
A large portion of the growing number of social media sites support some sort of human-directed tagging of posts.
Additionally, an increasing number of individuals are using tags as a form of query-based search to find information
(e.g., \citeNP{Diakopoulos2010}).

A growing body of research is focusing on the lifetime and growth characteristics of hashtags for real-time social networking sites such as Twitter
(e.g., \citeNP{Tsur2012, Chang2010}).
However, less is known about the underlying process when choosing particular tags during the creation of the tweet or post.
Modeling tag creation accurately (as opposed to modeling tag growth and behavior after creation) could allow a system to predict the hashtag that a user should use, given the contextual clues available.
There are numerous benefits to this type of tag prediction.
For example, users could be introduced to tags (and potentially communities surrounding those tags) that are of interest to them.
Also tag cleanup systems could be implemented where mistags are identified and corrected.

@@

\vspace{-1em}

@@

@@

During initial model testing, the attentional weighting for all words was equal, and set to $W/n$ for $n$ title words (and set similarly for body words).
The problem with this approach is that familiar stop words like ``the'' and ``?'', that are often observed across all tags have the same attentional weighting as highly predictive words like ``PHP''.
Granted the $S_{ji}$'s across all tags for context words like ``the'' are much lower (and more evenly distributed) than the few $S_{ji}$'s that spike for ``PHP'' (e.g., the tag \emph{PHP}).
However, this still seemed problematic for two reasons:
First, because stop words like ``the'' are observed with all tags and not highly predictive of any tag, one could argue that their presence is simply adding error in the model's tag activation calculation.
Second, because $W$ is held constant across posts, any attentional resources that are allocated to stop words are taken away from being used for highly predictive words.

@@

\subsubsection{Dataset}

\subsubsection{Distribution of Tag Frequency}
\emph{C\#} is the most often used tag, where about 11\% of posts were tagged with \emph{C\#}.
Tag frequency then drops off sharply.
The top 1\% of tags based on frequency of use account for approximately 60\% of total tag instances.
This is why accounting for tag frequency (by including ACT-R's base-level activation component) is so important for the StackOverflow dataset.

@@

Model tag predictions are based on the amount of word co-occurrences between words in the title and body of the post and tags used for that post.
To build the co-occurrence matrix between post words and tags, all title and body text and associated tags were extracted from the relational database.
The Python Natural Language Processing Toolkit \cite{Bird2009} was used to chunk (i.e., tokenize and lemmatize) the text in the post.
The tags were already chunked, but tag synonyms were converted to their root tag using the community-maintained StackOverflow tag synonym database.
Afterwords the processed word chunks and root tags were imported into a PostgreSQL table.

@@

\subsection{Current Model Issues}

Including the offset term in the model improved performance (see Figure \ref{fig:ROC}).
However by normalizing the top activations for all posts, the model cannot differentiate between a post where the top predicted tag is certain, and a post where the top-predicted tag is questionable.
One could argue that normalizing top activations for all posts should cause model performance to decrease, and the opposite was actually observed.
One hypothesis is that this normalization is required because the retrieval threshold of a tag is adjusted for each post, and changed so that only a small set of tags are retrieved for all posts.
Further research will explore this and other theories for the behavior of the offset component in the model.

@@

Alongside exploring the offset component of the model, our main planned research is to include the post author's past tagging behavior as an additional model predictor.
We are exploring a Bayesian updating technique that would adjust prior tag activations from global priors to user's priors as more tagging information about the user is collected.
For the StackOverflow dataset, adding user's tagging history to the model should drastically improve predictive power for veteran users,
since tags are tied to programming languages and specific technologies, and programmers tend to specialize on a few set of tags that are of interest to them.
This technique could also be deployed, since a post author must be signed into StackOverflow before creating a new post.
